---
title: "Lab 4 - Data Frame Basics"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```
Look at both the dataset file, and the file with its description, and answer the following questions:

What's the character delimiter? csv
Is there a row for column names? no
Are there any missing values? If so, how are they codified? No missing values
What is the data type of each column? 
  Sex=nominal
	Length=continuous
	Diameter=continuous
	Height=continuous
	Whole weight=continuous
	Shucked weight=continuous
	Viscera weight=continuous
	Shell weight=continuous
	Rings=integer
	
	
```{r}
	url <- "http://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data"
  abalone <- read.table(url, sep = ",")
  abalone <- read.table("abalone.data", sep = ",")
  
  # take a peek of first rows
  head(abalone)

  # take a peek of last rows
  tail(abalone)
  
  # check data frame's structure
  str(abalone, vec.len = 1)
  
  # vector of column names
  column_names <- c(
      'sex',
      'length',
      'diameter',
      'height',
      'whole_weight',
      'shucked_weight',
      'viscera_weight',
      'shell_weight',
      'rings'
  )

  # vector of data types (for each column)
  column_types <- c(
      'character',
      'real',
      'real',
      'real',
      'real',
      'real',
      'real',
      'real',
      'integer'   
  )
  
  abalone <- read.table(
      'abalone.data',
      col.names = column_names,
      colClasses = column_types,
      sep = ","
  )
  
  # check its structure again
  str(abalone, vec.len = 1)
```
	


	
	Your Turn:
	Read the Abalone data with the read.csv() function.
  Use the inputs col.names and colClasses to specify column names and their data types.
```{r}
  read.csv("abalone.data")
  abalone <- read.csv("abalone.data", col.names = column_names, colClasses = column_types, sep = ",")
```

Pittsburgh Bridges Data Set
  Are there column names? No
  What is the field separator? ","
  Are there any missing values? 
      Missing Attribute Values:
      Attribute #: 	# instances with missing values:
      2			        1
      6			        27
      7			        16
      8			        2
      9			        6
     10			        2
     11			        16
     12			        5
     13			        3
  What is the character for missing values (if any)?  "?"
  What is the data type of each variable (i.e. column)? 
       1.  IDENTIF	-	
       2.  RIVER	nomial
       3.  LOCATION	nomial
       4.  ERECTED	integer,nomial
       5.  PURPOSE	nomial
       6.  LENGTH	integer,nomial
       7.  LANES	integer,nomial
       8.  CLEAR-G	nomial
       9.  T-OR-D	nomial
       10. MATERIAL	nomial
       11. SPAN	nomial
       12. REL-L	nomial
       13. TYPE	nomial
  Download a copy of the data to your computer (use download.file()) and save it in a file named bridges.data
    download.file('http://archive.ics.uci.edu/ml/machine-learning-databases/bridges/bridges.data.version1', 'bridges.data')

```{r}
column_names <- c(
    'IDENTIF', 'RIVER', 'LOCATION', 'ERECTED', 'PURPOSE', 
    'LENGTH', 'LANES', 'CLEAR-G', 'T-OR-D', 'MATERIAL', 'SPAN', 'REL-L', 'TYPE'
)

column_types <-c(
    'character', 'character', 'integer', 'character', 'character', 'numeric', 'integer', 'character', 'character', 'numeric', 'character', 'character', 'character'
)

```


WARRIORS
```{r}
player<-c("Thompason","Curry","Green","Durant","Pachulia")
position<-c("SG","PG","PF","SF","C")
salary<-c(16663575, 12112359, 15330435, 26540100, 2898000)
points<-c(1742, 1999, 776, 1555, 426)
ppg<-c(22.3, 25.3, 10.2, 25.1, 6.1)
rookie<-c(FALSE, FALSE, FALSE, FALSE, FALSE)
warriors<-data.frame(player, position, salary, points, ppg, rookie)
typeof(warriors)
class(warriors)
warriors2<-list(player,position,salary,points,ppg, rookie)
warriors3<-data.frame(warriors2)
binded<-cbind(player, position,salary,points,ppg,rookie)
warriors_binded<-data.frame(binded)

# download csv file into your working directory
csv <- "https://github.com/ucb-stat133/stat133-fall-2017/raw/master/data/nba2017-players.csv"
download.file(url = csv, destfile = 'nba2017-players.csv')
dat<-read.csv('nba2017-players.csv', stringsAsFactors = FALSE)
```
```{r}
#Your turn:
#Use bracket notation, the dollar operator, as well as concepts of logical subsetting and indexing to:

#Display the last 5 rows of the data.
  tail(dat)
#Display those rows associated to players having height less than 70 inches tall.
  dat$player[dat$height < 70]
#Of those players that are centers (position C), display their names and salaries.
  dat[dat$height<70 && dat$position == 'C', c(1,9)]
#Create a data frame durant with Kevin Durant's information (i.e. row).
  durant <- dat[dat$player == "Kevin Durant",]
  durant
#Create a data frame ucla with the data of players from college UCLA ("University of California, Los Angeles").
  ucla <- dat[dat$college == "University of California, Los Angeles",]
  ucla
#Create a data frame rookies with those players with 0 years of experience.
  rookies <- dat[dat$experience == 0,]
  rookies
#Create a data frame rookie_centers with the data of Center rookie players.
rookie_centers <- rookies[rookies$position == "C", ]
rookie_centers
#Create a data frame top_players for players with more than 50 games and more than 100 minutes played.
top_players <- dat[dat$games > 50 && dat$minutes >= 100]
top_players
#What's the largest height value?
max(dat$height)

#What's the minimum height value?
min(dat$height)
#What's the overall average height?
mean(dat$height)
#Who is the tallest player?
dat$player[dat$height == max(dat$height)]
#Who is the shortest player?
dat$player[dat$height == min(dat$height)]
#Which are the unique teams?

#How many different teams?

#Who is the oldest player?
dat$player[dat$age == max(dat$age)]
#What is the median salary of all players?
median(dat$salary)
#What is the median salary of the players with 10 years of experience or more?
median(dat$salary[dat$experience >= 10])
#What is the median salary of Shooting Guards (SG) and Point Guards (PG)?
median(dat$salary[dat$position == "SG" || dat$position == "PG"])
#What is the median salary of Power Forwards (PF), 29 years or older, and 74 inches tall or less?
median(dat$salary[dat$position == "PF" && dat$age >= 29 && dat$height <= 74])
#How many players scored 4 points or less?
sum(dat$points <= 4)
#Who are those players who scored 4 points or less?
dat$player[dat$points <= 4]
#Who is the player with 0 points?
dat$player[dat$points == 0]
#How many players are from "University of California, Berkeley"?
sum(dat$college == "University of California, Berkeley")
#Are there any players from "University of Notre Dame"? If so how many and who are they?
sum(dat$college == "University of Notre Dame")
dat$player[dat$college == "University of Notre Dame"]
#Are there any players with weight greater than 260 pounds? If so how many and who are they?
sum(dat$weight > 260)
dat$player[dat$weight > 260]
#How many players did not attend a college in the US?

#Who is the player with the maximum rate of points per minute?
#dat$player[dat$points == max(points)]
#Who is the player with the maximum rate of three-points per minute?

#Who is the player with the maximum rate of two-points per minute?

#Who is the player with the maximum rate of one-points (free-throws) per minute?

#Create a data frame gsw with the name, height, weight of Golden State Warriors (GSW)
gsw <- dat[dat$team == "GSW", c(1,4,5)]
gsw
#Display the data in gsw sorted by height in increasing order (hint: see ?sort and ?order)
order(gsw, decreasing = FALSE)
#Display the data in gsw by weight in decreasing order (hint: see ?sort and ?order)

#Display the player name, team, and salary, of the top 5 highest-paid players (hint: see ?sort and ?order)

#Display the player name, team, and points3, of the top 10 three-point players (hint: see ?sort and ?order)

```

